Irt plot extra

### 2022-11-14 manuscript_v2.Rmd
-gtExtra figure conflit resolved -full pdf for round 2 updated.

keywords          : "light exposure, light-related behaviours, non-visual effects of light, psychometrics"
wordcount         : "6358"

@@

An extensive body of scientific evidence suggests that the imbalance of light and dark exposure disrupts humans’ light-dependent physiological systems [@Lunn.2017]. Subsequently, this disruption gives rise to a series of adverse health consequences, including the alteration of hormonal rhythms, increased cancer rates, cardiovascular diseases, and metabolic disorders, such as obesity and type II diabetes [@Lunn.2017; @navara2007dark; @Chellappa.2019]. These findings have sparked a significant call for assessment and guidance regarding healthy light exposure as exemplified by a recently published set of consensus-based experts’ recommendations with specific requirements for indoor light environments during the daytime, evening, and nighttime [@Brown.2022]. Furthermore, building on earlier attempts [e.g. @hubalek_ambulant_2006], there was a recent push toward the development and use of portable light loggers to improve ambulant light assessment and gain more insight into the NIF effects of light on human health in field conditions [@stampfli_light-dosimeter_2021; @van_duijnhoven_recommendations_2017; @webler_towards_2021; @aarts2017performance]. Attached to different body parts (e.g., wrist; head, at eye level; chest), these devices allow for the objective measurement of individual photic exposure patterns under real-world conditions and thus are a valuable tools for field studies.. Nevertheless, these devices also encompass limiting factors such as potentially being intrusive (e.g., when eye-level worn), yielding the risk of getting covered (e.g., when wrist- or chest-worn) and requiring (monetary) resources and expertise for acquisition and maintenance of the devices.

@@

To date, little effort has been made to understand and capture these activities. **Supplementary Table 1** summarises the existing questionnaire literature assessing light exposure-related properties. However, only a few questions of these existing tools were associated with light exposure-related behaviour. For example, the “Munich Chronotype  Questionnaire” [@roenneberg2003life], a popular self-report tool for identifying chronotypes via mid-sleep times, includes questions about the individual’s typical time spent outdoors on workdays and free days. The Visual Light Sensitivity Questionnaire-8 [@verriotto2017new] and Photosensitivity Assessment Questionnaire (@bossini2006sensibilita) are a couple of self-report tools measuring visual light sensitivity. They contain single items which probe the preference for specific light situations such as: “In the past month, how often did you need to wear dark glasses on cloudy days or indoors?” [@verriotto2017new];"I prefer rooms that are in semi-darkness."; [@bossini2006sensibilita]. In addition, the “Pittsburgh Sleep Quality Index” [@buysse1989pittsburgh], is a popular measure of sleep quality. It contains questions about bedtime and wake-up times, which are relevant to light exposure around bedtime. However, none of these questionnaires provides a scalable solution to capture light exposure-related behaviour in various lighting situations. To fill this gap, we here present the development process of a novel self-reported tool - the “Light Exposure Behaviour Assessment” (LEBA) - for characterizing diverse light exposure-related behaviours.

@@

For reliability estimation, the "psych" package was applied [@R-psych]. Though Cronbach's internal consistency coefficient alpha is widely used for estimating internal consistency, it tends to deflate the estimates for Likert-type data since the calculation is based on the Pearson-correlation matrix, which requires response data to be continuous in nature [@gadermann2012estimating; @zumbo2007ordinal]. Subsequently, we reported ordinal alpha for each factor obtained in the EFA which was suggested as a better reliability estimates or ordinal data [@zumbo2007ordinal]. We also estimated the internal consistency reliability of the total scale using McDonald's $\omega_t$ coefficient, which was suggested as a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. Both ordinal alpha and McDonald's $\omega_t$ coefficient values range between 0 to 1, where higher values represent better reliability.

@@

full.tab.1 %>% 
# histogram and density plots
  gtExtras::gt_plt_dist(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_plt_dist(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "grey",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("Summary Descriptives (n=690)"),
    subtitle = md("Items 01-24 ")) %>% 
#create groups of columns
tab_spanner(
  label = "Item",
  columns = Item:Stem
) %>%
  cols_align(
    align = "left",
    columns = Item:Stem
  ) %>% 
  tab_spanner(
    label = "Summary Statistics",
    columns = mean:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%

     tab_footnote(
     footnote = md("**Shapiro–Wilk test**"),
     locations = cells_column_labels(columns = `S-W Statistics`)
  ) %>% 
# change column names to appear in the table
cols_label(
  Item = ("Items"),
  Stem = ("Stem"),
  mean = ("Mean"),
  sd = (("SD")),
  `S-W Statistics` = ("SW"),
  hist_data = "Histogram",
  dens_data = "Density"
)  %>% 
# set alignment as per wish
  cols_align(
    align = "center",
    columns = mean:Always
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
  gt_plt_dot(
    mean,
    Item,
    palette = "ggthemes::fivethirtyeight"
  ) %>% tab_options(
    table.font.size = px(14L)) %>% 
  cols_width(
    Stem ~ px(200)) %>% 
 gtsave("Figures/Figure2.png", vwidth = 6000)

full.tab.2 %>% 
# histogram and density plots
  gtExtras::gt_plt_dist(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_plt_dist(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "grey",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("Summary Descriptives (n=690)"),
    subtitle = md("Items 25-48 ")) %>% 
#create groups of columns
tab_spanner(
  label = "Item",
  columns = Item:Stem
) %>%
  cols_align(
    align = "left",
    columns = Item:Stem
  ) %>% 
  tab_spanner(
    label = "Summary Statistics",
    columns = mean:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%

@@

Applying varimax rotation, we conducted three rounds of EFA with the initial 48 items and gradually discarded problematic items (cross-loading items and items with factor loading \<.30).
Finally, a five-factor EFA solution with 25 items was accepted with all factor-loading higher than .30 and no cross-loading greater than .30. Table \@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute values of the factor-loadings ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)`, indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` and `r apa(max.com, 2, F)`. However, the histogram of the absolute values of nonredundant residual correlations (Figure \@ref(fig:efa-plot-print)-D) displayed that `r apa(sumR,0, T)`% of correlations were greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018]. Subsequently, we fitted a six-factor solution, where a factor with only two salient variables emerged, thus disqualifying the six-factor solution (**Supplementary Table 4**).

@@

cfa.gt %>% 
gt() %>% 
# histogram and density plots
  gtExtras::gt_plt_dist(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_plt_dist(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "#E64B3599",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("**Light Exposure Behaviour Assessment**"),
    subtitle = md("Summary Descriptives CFA Sample (n=262)")
  ) %>% 
  # tab_footnote(
  #   footnote = md("**Histogram**"),
  #   locations = cells_column_labels(columns = hist_data)
  # ) %>%
  # tab_footnote(
  #   footnote = md("**Density**"),
  #   locations = cells_column_labels(columns = dens_data)
  # ) %>% 
#create groups of columns
tab_spanner(
  label = "Items",
  columns = Items
) %>%
  tab_spanner(
    label = "Summary Statistics",
    columns = nr:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%
# change column names to appear in the table
cols_label(
  Items = (""),
  nr = ("n"),
  mean = ("Mean"),
  med = ("Median "),
  sd = (("SD")),
  hist_data = "Histogram",
  dens_data = "Density"
) %>%
# set alignment as per wish
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
  gt_plt_dot(
    mean,
    Items,
    palette = "ggthemes::fivethirtyeight") %>% 
  tab_row_group(
      label = md("**F5: Using light in the morning and during daytime**"),
      rows = c(1,3,12,13,22,23) 
        ) %>% 
  tab_row_group(
      label = md("**F4: Using light before bedtime**"),
      rows = c(15,16,17,19) 
        ) %>% 
  tab_row_group(
      label = md("**F3: Using phone and smartwatch in bed**"),
      rows = c(2,13,14,20,21) 
        ) %>% 
  tab_row_group(
      label = md("**F2: Spending time outdoors**"),
      rows = c(4,5,6,7,8,9) 
        ) %>% 
  tab_row_group(
      label = md("**F1: Wearing blue light filters**"),
      rows = c(10,11,18) 
        ) %>% 
  gtsave("Figures/S2_Fig.png",vwidth = 7000)

@@

##IRT reliability (marginal reliability of scales)
#F1.marginal.rel <- marginal_rxx(F1_fit)
#F2.marginal.rel <-marginal_rxx(F2_fit.redu)
#F3.marginal.rel <-marginal_rxx(F3_fit.redu)
#F4.marginal.rel <-marginal_rxx(F4_fit.redu)
#F5.marginal.rel <-marginal_rxx(F5_fit)

>>>>

keywords          : "light exposure, light-related behaviours, non-visual effects of light, psychometrics"
wordcount         : "6249"

@@

An extensive body of scientific evidence suggests that the imbalance of light and dark exposure disrupts humans’ light-dependent physiological systems [@Lunn.2017]. Subsequently, this disruption gives rise to a series of adverse health consequences, including the alteration of hormonal rhythms, increased cancer rates, cardiovascular diseases, and metabolic disorders, such as obesity and type II diabetes [@Lunn.2017; @navara2007dark; @Chellappa.2019]. These findings have sparked a significant call for assessment and guidance regarding healthy light exposure as exemplified by a recently published set of consensus-based experts’ recommendations with specific requirements for indoor light environments during the daytime, evening, and nighttime [@Brown.2022]. Furthermore, building on earlier attempts [e.g. @hubalek_ambulant_2006], there was a recent push toward the development and use of portable light loggers to improve ambulant light assessment and gain more insight into the NIF effects of light on human health in field conditions [@stampfli_light-dosimeter_2021; @van_duijnhoven_recommendations_2017; @webler_towards_2021; @aarts2017performance]. Attached to different body parts (e.g., wrist; head, at eye level; chest), these devices allow for the objective measurement of individual photic exposure patterns under real-world conditions and thus are valuable tools for field studies. Nevertheless, these devices also encompass limiting factors such as potentially being intrusive (e.g., when eye-level worn), yielding the risk of getting covered (e.g., when wrist- or chest-worn) and requiring (monetary) resources and expertise for acquisition and maintenance of the devices.

@@

To date, little effort has been made to understand and capture these activities. **Supplementary Table 1** summarises the existing questionnaire literature assessing light exposure-related properties. However, only a few questions of these existing tools were associated with light exposure-related behaviour. For example, the “Munich Chronotype  Questionnaire” [@roenneberg2003life], a popular self-report tool for identifying chronotypes via mid-sleep times, includes questions about the individual’s typical time spent outdoors on workdays and free days. The Visual Light Sensitivity Questionnaire-8 [@verriotto2017new] and Photosensitivity Assessment Questionnaire [@bossini2006sensibilita] are a couple of self-report tools measuring visual light sensitivity. They contain single items which probe the preference for specific light situations such as: “In the past month, how often did you need to wear dark glasses on cloudy days or indoors?” [@verriotto2017new];"I prefer rooms that are in semi-darkness."; [@bossini2006sensibilita]. In addition, the “Pittsburgh Sleep Quality Index” [@buysse1989pittsburgh], is a popular measure of sleep quality. It contains questions about bedtime and wake-up times, which are relevant to light exposure around bedtime. However, none of these questionnaires provides a scalable solution to capture light exposure-related behaviour in various lighting situations. To fill this gap, we here present the development process of a novel self-reported tool - the “Light Exposure Behaviour Assessment” (LEBA) - for characterizing diverse light exposure-related behaviours.

@@

For reliability estimation, the "psych" package was applied [@R-psych]. Though Cronbach's internal consistency coefficient alpha is widely used for estimating internal consistency, it tends to deflate the estimates for Likert-type data since the calculation is based on the Pearson-correlation matrix, which requires response data to be continuous in nature [@gadermann2012estimating; @zumbo2007ordinal]. Subsequently, we reported ordinal alpha for each factor obtained in the EFA which was suggested as a better reliability estimates for ordinal data [@zumbo2007ordinal]. We also estimated the internal consistency reliability of the total scale using McDonald's $\omega_t$ coefficient, which was suggested as a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. Both ordinal alpha and McDonald's $\omega_t$ coefficient values range between 0 to 1, where higher values represent better reliability.

@@

full.tab.1 %>% 
# histogram and density plots
  gtExtras::gt_plt_dist(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_plt_dist(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "grey",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("Summary Descriptives (n=690)"),
    subtitle = md("Items 01-24 ")) %>%
  cols_align(
    align = "left",
    columns = Item:Stem
  ) %>% 
  tab_spanner(
    label = "Summary Statistics",
    columns = mean:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%
     tab_footnote(
     footnote = md("**Shapiro–Wilk test**"),
     locations = cells_column_labels(columns = `S-W Statistics`)
  ) %>% 
# change column names to appear in the table
cols_label(
  Item = ("Items"),
  Stem = ("Stem"),
  mean = ("Mean"),
  sd = (("SD")),
  `S-W Statistics` = ("SW"),
  hist_data = "Histogram",
  dens_data = "Density"
)  %>% 
# set alignment as per wish
  cols_align(
    align = "center",
    columns = mean:Always
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
  gt_plt_dot(
    mean,
    Item,
    palette = "ggthemes::fivethirtyeight"
  ) %>% tab_options(
    table.font.size = px(14L)) %>% 
  cols_width(
    Stem ~ px(200)) %>% 
 gtsave("Figures/Figure2.png", vwidth = 6000)

full.tab.2 %>% 
# histogram and density plots
  gtExtras::gt_plt_dist(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_plt_dist(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "grey",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("Summary Descriptives (n=690)"),
    subtitle = md("Items 25-48 ")) %>% 
#create groups of columns
# tab_spanner(
#   label = "Item",
#   columns = Item:Stem
# ) %>%
  cols_align(
    align = "left",
    columns = Item:Stem
  ) %>% 
  tab_spanner(
    label = "Summary Statistics",
    columns = mean:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%

@@

Applying varimax rotation, we conducted three rounds of EFA with the initial 48 items and gradually discarded problematic items (cross-loading items and items with factor loading <.30).
Finally, a five-factor EFA solution with 25 items was accepted with all factor-loading higher than .30 and no cross-loading greater than .30. Table \@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute values of the factor-loadings ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)`, indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` and `r apa(max.com, 2, F)`. However, the histogram of the absolute values of nonredundant residual correlations (Figure \@ref(fig:efa-plot-print)-D) displayed that `r sumR`% of correlations were greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018]. Subsequently, we fitted a six-factor solution, where a factor with only two salient variables emerged, thus disqualifying the six-factor solution (**Supplementary Table 4**).

@@

cfa.gt %>% 
gt() %>% 
# histogram and density plots
  gtExtras::gt_plt_dist(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_plt_dist(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "#E64B3599",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("**Light Exposure Behaviour Assessment**"),
    subtitle = md("Summary Descriptives CFA Sample (n=262)")
  ) %>% 
  # tab_footnote(
  #   footnote = md("**Histogram**"),
  #   locations = cells_column_labels(columns = hist_data)
  # ) %>%
  # tab_footnote(
  #   footnote = md("**Density**"),
  #   locations = cells_column_labels(columns = dens_data)
  # ) %>% 
#create groups of columns
# tab_spanner(
#   label = "Items",
#   columns = Items
# ) %>%
  tab_spanner(
    label = "Summary Statistics",
    columns = nr:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%
# change column names to appear in the table
cols_label(
  Items = (""),
  nr = ("n"),
  mean = ("Mean"),
  med = ("Median "),
  sd = (("SD")),
  hist_data = "Histogram",
  dens_data = "Density"
) %>%
# set alignment as per wish
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
gt_plt_dot(
  mean,
  Items,
  palette = "ggthemes::fivethirtyeight") %>%
  tab_row_group(
      label = md("**F5: Using light in the morning and during daytime**"),
      rows = c(1,3,12,13,22,23) 
        ) %>% 
  tab_row_group(
      label = md("**F4: Using light before bedtime**"),
      rows = c(15,16,17,19) 
        ) %>% 
  tab_row_group(
      label = md("**F3: Using phone and smartwatch in bed**"),
      rows = c(2,13,14,20,21) 
        ) %>% 
  tab_row_group(
      label = md("**F2: Spending time outdoors**"),
      rows = c(4,5,6,7,8,9) 
        ) %>% 
  tab_row_group(
      label = md("**F1: Wearing blue light filters**"),
      rows = c(10,11,18) 
        ) %>% 
  gtsave("Figures/S2_Fig.png",vwidth = 7000)

@@

##IRT reliability (marginal reliability of scales)
# F1.marginal.rel <- marginal_rxx(F1_fit)
# F2.marginal.rel <-marginal_rxx(F2_fit.redu)
# F3.marginal.rel <-marginal_rxx(F3_fit.redu)
# F4.marginal.rel <-marginal_rxx(F4_fit.redu)
# F5.marginal.rel <-marginal_rxx(F5_fit)
